For Induction-Induction, Induction is Enough

### 2019-11-12 paper/paper.tex
more related work

Forsberg \cite{forsberg-phd} specifies inductive-inductive types and develops a
semantics based on dialgebras. He also provides a reduction of IITs to indexed
inductive types, but only constructs a weak version of elimination.

@@

Streicher \cite{streichersemantics} presents an interpretation of the well-formed
presyntax of a type theory into a categorical model, which is an important
ingredient in constructing an initial model, although he does not present
details on the construction of the term model or its initiality proof. Our paper an indexed variant of his construction (see Subsection \ref{sec:streicher} for a comparison).

Voevodsky was interested in constructing initial models of type
theories from presyntaxes. Inspired by this, Brunerie \cite{brunerie}
formalised Streicher's proof in Agda for a type theory with $\Pi$,
$\Sigma$, $\mathbb{N}$, identity types and an infinite hierarchy of
universes. He used UIP, function extensionality and quotient types in
his formalisation. Our proof uses a similar approach, but we construct
a much simpler type theory which does not have any computation rules,
hence we can avoid using quotients.

Intrinsic (well-typed) syntax for type theories were constructed
before using IITs \cite{chapman09eatitself}, inductive-recursive types
\cite{nisse,Altenkirch:2014:CO:2631172.2631176} and QIITs
\cite{ttintt}. In this paper we avoid using such general classes of
inductive types, as our goal is to reduce IITs to indexed inductive
types.

Reducing general classes of inductive types to simpler classes has a
long tradition in type theory. Indexed W-types were reduced to W-types
\cite{indexedcont} (using the essentially Streicher's idea of preterms
and a typing predicate), small inductive-recursive types to indexed
W-types \cite{malatasta13smallir}, mutual inductive types to indexed
W-types \cite{mutual}. Quotient inductive-inductive types (and
inductive-inductive types) can be reduced to quotient inductive types
using the reduction of generalised algebraic theories to essentially
algebraic theories \cite{gat}.

>>>>

Forsberg \cite{forsberg-phd} specifies inductive-inductive types and develops a
semantics based on dialgebras. He also provides a reduction of IITs to indexed
inductive types, but only constructs a weak version of the eliminator.

@@

Streicher \cite{streichersemantics} presents an interpretation of the well-formed
presyntax of a type theory into a categorical model, which is an important
ingredient in constructing an initial model, although he does not present
details on the construction of the term model or its initiality proof. Our initiality proof can be seen as an indexed variant of his construction (see Subsection \ref{sec:streicher} for a comparison).

Voevodsky was interested in constructing initial models of type
theories from presyntaxes. Inspired by this, Brunerie \cite{brunerie}
formalised Streicher's proof in Agda for a type theory with $\Pi$,
$\Sigma$, $\mathbb{N}$, identity types and an infinite hierarchy of
universes. He used UIP, function extensionality and quotient types in
his formalisation. In this paper we construct a type theory without
computation rules, hence we avoid using quotients.

Intrinsic (well-typed) syntax for type theories were constructed using
IITs \cite{chapman09eatitself}, inductive-recursive types
\cite{nisse,Altenkirch:2014:CO:2631172.2631176} and QIITs
\cite{ttintt}. In this paper we avoid using such general classes of
inductive types as our goal is to reduce IITs to indexed inductive
types.

Reducing general classes of inductive types to simpler classes has a
long tradition in type theory. Indexed W-types were reduced to W-types
\cite{indexedcont} (using the essentially Streicher's idea of preterms
and a typing predicate), small inductive-recursive types to indexed
W-types \cite{malatasta13smallir}, mutual inductive types to indexed
W-types \cite{mutual}, W-types to natural numbers and quotients
\cite{Ahrens2019}. Quotient inductive-inductive types (and
inductive-inductive types) can be reduced to quotient inductive types
using the reduction of generalised algebraic theories to essentially
algebraic theories \cite{gat}.
